Data

### 2019-02-18 analysis/data.Rmd
data description

Here we describe the
* raw data, i.e. the Reference Trajectory data set
* pre-processed data, i.e. the extracted arrivals and the transition points.

>>>>

# Data Files

We have two directories for data

1. `data-raw`: it contains the raw source data. For this project these are the Reference Trajectory
   files as from Google Cloud Platform. See https://github.com/euctrl-pru/reftrj .
   Due to their big size these are not duplicated/version controlled on this repo.
1. `data`: here are all the files necessary for the analysis as performed in the paper.

## From `data-raw` to `data`

The raw data filenames are as follows:

* flight information, i.e. `cpr_fr24_flights_2017-08-01.csv.bz2`
* positions data, i.e. `mas_05_cpr_fr24_synth_positions_2017-08-01.csv.bz2`

PRU's [Open Reference Trajectory dataset](https://github.com/euctrl-pru/reftrj)
makes available the files needed for the analysis in this project,
you can retrieve them either by saving from your browser the relevant linked
flight and position files or use Google `gsutil` as follows:

```{bash ,eval=FALSE}
# enable relevant Python virtual environment (Python 2) if you have one, and you should.
$ pyenv activate anaconda2-4.2.0
$ cd data-raw
$ gsutil cp \
    gs://pru-trajectories/sources/merged/overnight_cpr_fr24/raw_cpr_fr24_positions_2017-08-08.csv.bz2 .
```

As explained in the paper, you can use `R/extract-arrivals.R` to get to CSV files
containing only the flights and positions for your airport of initerest.
You can edit the relevant configuration values at the beginning of the script before launching it.
Yuo will get:

* Flights details for arrivals at `<airport>` on `<date>`
  are in a CSV file `<airport>_flt_rt_<date>.csv`.
  For example `egll_flt_rt_2017-08-01.csv` is for arrivals to
  London Heathrow Airport (ICAO's id EGLL) on Aug, 1 2017.
  Key data items:
    * flight_id     unique flight identification
    * callsign      callsign flight identification of flight
    * aircraft_reg  aircraft registration
    * aircraft_type aircraft type, ICAO coding (e.g. A388, B772)
    * adep / ades   aerodrome of departure / destination

* Position reports for arrival at `<airport>` on `<date>`
  are in a CSV file `<airport>_pos_rt_<date>.csv`.
  For example `eidw_pos_rt_2017-08-01.csv` is for arrivals to
  Dublin Airport (ICAO's id EIDW) on Aug, 1 2017.

The focus of this project is for the terminal area, so we have further filtered the data
and kept the last 40 NM from the airport reference position (ARP).

The position reports within 40 nautical miles for arrival at
`<airport>` on `<date>`  are in a CSV file 
`<airport>_pos_rt_40NM_<date>.csv`.
These are a subset from `<airport>_pos_rt_<date>.csv`.
For example `eidw_pos_rt_40NM_2017-08-02.csv` is for arrivals'
positions within 40 nautical miles to Dublin Airport
(ICAO's id EIDW) on Aug, 2 2017.
Key data items:
  * flight_id     unique flight identification
  * timestamp     position report time
  * latitude      latitude coordinate (decimal notation)
  * longitude     longitude coordinate (decimal notation)
  * on_ground     on-ground / aloft flag
  * distance      [not sure ~~] miles flown [?]
  * distance_arp  distance position to aerodrome reference point

Further we have extracted the transition points from the holding/point merge areas
using the script `R/extract-transition-points.R` which produces
the files `<airport>_transition_points_<date>.csv`.

## Version controlled data

We version control only the following data sets in `data`

* `<airport>_flt_rt_<date>.csv`
* `<airport>_pos_rt_40NM_<date>.csv`
* `<airport>_transition_points_<date>.csv`
